---
title: "Chapter 1 Excercises"
author: "S.D Penning"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(jtools)

theme_set(theme_classic())
```

# 1.19 GPA

## Observations:

1. n = 120
2. no replication
3. selection at random
4. all from the same college (random effect) if we want to know about all colleges but fixed effect if we want to know about this college

```{r}
gpa.df = read.csv(".\\data\\Chapter 1 data sets\\CH01PR19.txt", header = FALSE,  strip.white=TRUE, sep = "", col.names = c("GPA","ACT"))
```

```{r}
ggplot(data=gpa.df, aes(x=ACT, y=GPA,colour=GPA)) +
  geom_point() +
  ggtitle("GPA vs. ACT scores")
  
```

## a) Least squares estimates of $\beta_{0}$ and $\beta_{1}$
```{r}
library(moderndive)
gpa.lm = lm(data=gpa.df, formula = GPA ~ ACT)

reg_params <- get_regression_table(gpa.lm)
reg_params
```

$\beta_{0}=$ `r reg_params$estimate[1]`, confidence interval: (`r reg_params$lower_ci[1]`, `r reg_params$upper_ci[1]`)

$\beta_{1}=$ `r reg_params$estimate[2]`, confidence interval: (`r reg_params$lower_ci[2]`, `r reg_params$upper_ci[2]`)


```{r}
gpa.lm.diag.metrics <- augment(gpa.lm)

gpa.lm.summary = summary(gpa.lm)
```


## b) Plot the estimated regression function and the data

```{r}
ggplot(data=gpa.lm.diag.metrics, aes(x=ACT, y=GPA)) +
  geom_point() +
  geom_line(aes(y=.fitted))
```

```{r}
ggplot(gpa.lm.diag.metrics, aes(ACT, GPA)) +
  geom_point() +
  stat_smooth(method = lm, se = TRUE) +
  geom_segment(aes(xend = ACT, yend = .fitted), color = "red", size = 0.3)
```

The model does not wit the data particularity well. The model only explains `r round(gpa.lm.summary$adj.r.squared*100,1)` % of the variation in the data

## c) Obtain a point estimate for ACT = 30

```{r}
new_data = data.frame(ACT=30)
gpa.pred = predict(gpa.lm, newdata = new_data, interval = "prediction")

```

GPA(30) = `r gpa.pred[1]`. Prediction interval = (`r gpa.pred[2]`, `r gpa.pred[3]`)

## d) What is the point estimate of the change in mean reposnse when the test score increases by 1 point

$\beta_{1}=$ `r reg_params$estimate[2]`, confidence interval: (`r reg_params$lower_ci[2]`, `r reg_params$upper_ci[2]`)


```{r}
library(ggfortify)
autoplot(gpa.lm)
```
```{r}
# Cook's distance
plot(gpa.lm, 4)
# Residuals vs Leverage
plot(gpa.lm, 5)
```
### Remove the outliers

```{r}
library(dplyr)
library(lmtest)

gpa.df2 <- gpa.df %>%
  slice(-c(2,9,115))

gpa.lm2 = lm(data=gpa.df2, formula = GPA ~ ACT)

bptest(gpa.lm2)
```

```{r}

reg_params2 <- get_regression_table(gpa.lm2)
reg_params2

autoplot(gpa.lm2)
summary(gpa.lm2)
summary(gpa.lm)

gpa.lm.diag.metrics2 <- augment(gpa.lm2)
ggplot(gpa.lm.diag.metrics2, aes(ACT, GPA)) +
  geom_point() +
  stat_smooth(method = lm, se = TRUE) +
  geom_segment(aes(xend = ACT, yend = .fitted), color = "red", size = 0.3)

```

