---
title: "Chapter1-1_24"
author: "S.D Penning"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      theme: united
      highlight: tango
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

# 1.20 Copier Mainternance

```{r}
maint.dat = read.csv(".\\data\\Chapter 1 data sets\\CH01PR20.txt", header = FALSE,  strip.white=TRUE, sep = "", col.names = c("Minutes","Services"))
```

```{r}
ggplot(data=maint.dat, aes(x=Services,y=Minutes)) +
  geom_point() +
  ggtitle("Maintenance time vs. Number of Services")
```

```{r}
maint.lm = lm(data = maint.dat, formula = Minutes ~ -1 + Services)
summary(maint.lm)

```
```{r}
library(moderndive)

reg_params <- get_regression_table(maint.lm)
reg_params
```

# a) Estimated regression function


$\beta_{0}=0$

$\beta_{1}=$ `r reg_params$estimate[1]`, confidence interval: (`r reg_params$lower_ci[1]`, `r reg_params$upper_ci[1]`)

```{r}
maint.aug = augment(maint.lm, data = maint.dat, se_fit = TRUE, interval = c("prediction"))

```

# b) Plot estimated regression function

```{r}
ggplot(data=maint.aug, aes(x=Services,y=Minutes)) +
  geom_point() +
  ggtitle("Maintenance time vs. Number of Services") +
  geom_line(aes(y=.fitted))+
  geom_line(aes(y=.fitted+2*.se.fit)) +
  geom_line(aes(y=.fitted-2*.se.fit))

```

# c) Interpret $\beta_{0}$

$\beta_{0}$ must be zero in this case. Negative values for zero services do not make sense.

```{r}
Services = c(5)
maint.new = as_tibble(Services)
maint.new.pred = predict.lm(object = maint.lm, newdata = maint.new, se.fit = TRUE)
```

# d) Point estimate when X=5 copiers

`r maint.new.pred$fit` , s.e = `r maint.new.pred$se.fit`


# 1.24

```{r}
sum_resid = sum(maint.aug$.resid)

sum_resid_sq = sum(maint.aug$.resid^2)

MSE = sum_resid_sq/(length(maint.dat$Minutes)-1)

se = sqrt(MSE)
```


$\sum{\varepsilon_{i}^{2}}$ = `r sum(maint.aug$.resid^2)`


$\sum{\varepsilon_{i}^{2}}$ = Min(Q)

MSE = `r MSE`

se = `r se` minutes


