---
title: "Chapter1-1_21"
author: "S.D Penning"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
library(broom)
library(moderndive)
```

# 1.21 Airfreight breakage

```{r}
break.dat = read.csv(".\\data\\Chapter 1 data sets\\CH01PR21.txt", header = FALSE,  strip.white=TRUE, sep = "", col.names = c("Breakages","Transfers"))
```

## Inspect the data

### Check the types and range

```{r}
glimpse(break.dat)
```

## Check for missing values

```{r}
skim(break.dat)
```

## Compute correlation values
```{r}

break.dat %>% 
  get_correlation(formula = Breakages ~ Transfers)

break.dat %>%
  summarise(correlation = cor(Breakages, Transfers))
```

## Plot the raw data

```{r}
ggplot(data = break.dat, aes(x=Transfers, y=Breakages))+
  geom_point() +
  geom_jitter(width = 0.05) +
  geom_smooth(method = "lm", se = FALSE)
```

# a) Obtain the estimated regression function

```{r}
break.lm = lm(data = break.dat, formula = Breakages ~ Transfers)
summary(break.lm)
```
```{r}
library(moderndive)

reg_params <- get_regression_table(break.lm)
reg_params
```
$\beta_{0}=$ `r reg_params$estimate[1]`, confidence interval: (`r reg_params$lower_ci[1]`, `r reg_params$upper_ci[1]`)

$\beta_{1}=$ `r reg_params$estimate[2]`, confidence interval: (`r reg_params$lower_ci[2]`, `r reg_params$upper_ci[2]`)

```{r}
Transfers = c(1)
break.new = as_tibble(Transfers)
break.new.pred = predict.lm(object = break.lm, newdata = break.new, se.fit = TRUE)
b_1 = break.new.pred$fit
```

# b) Point estimate when X=1 transfers

`r break.new.pred$fit` , s.e = `r break.new.pred$se.fit`

# c) Estimated increase from 1 to 2 transfers

```{r}
Transfers = c(1,2)
break.new = as_tibble(Transfers)
break.new.pred = predict.lm(object = break.lm, newdata = break.new, se.fit = TRUE)
delta = break.new.pred$fit[2] - break.new.pred$fit[1]
se = sqrt(break.new.pred$se.fit[1]^2+break.new.pred$se.fit[2]^2)
```

Increase = `r delta` , s.e = `r se`

# d) Verify that the regression line goes through $(\bar{X},\bar{Y})$

```{r}
B_Bar = mean(break.dat$Breakages)
T_Bar = mean(break.dat$Transfers)
```

```{r}
Transfers = c(T_Bar)
break.new = as_tibble(Transfers)
break.new.pred = predict.lm(object = break.lm, newdata = break.new, se.fit = TRUE)
equality = (b_1 == break.new.pred$fit[1])
```

The regression line goes through $(\bar{X},\bar{Y})$ is `r equality`
