---
title: "Chapter1-1_22"
author: "S.D Penning"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      theme: united
      highlight: tango
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
library(broom)
library(moderndive)
library(visdat)
```

```{r}
hard.dat = read.csv(".\\data\\Chapter 1 data sets\\CH01PR22.txt", header = FALSE, sep = "", col.names = c('Hardness', 'Time'))
```


# Check for missing data

```{r}
vis_miss(hard.dat)
```

No missing data

# Visualise the data

```{r}
ggplot(data=hard.dat, aes(x=Time, y=Hardness)) +
  geom_point() +
  ggtitle(label = "Scatterplot of Hardness vs. Setting time")
```

# Obtain the estimated regression function

```{r}
hard.lm = lm(data=hard.dat, formula = Hardness ~ Time, qr = TRUE)
summary(hard.lm)
```
The fit appears to be quite good with high adjusted R squared and low RSE

```{r}
reg_params <- get_regression_table(hard.lm)
reg_params  
```

$\beta_{0}=$ `r reg_params$estimate[1]`, confidence interval: (`r reg_params$lower_ci[1]`, `r reg_params$upper_ci[1]`)

$\beta_{1}=$ `r reg_params$estimate[2]`, confidence interval: (`r reg_params$lower_ci[2]`, `r reg_params$upper_ci[2]`)

# Obtain point estimate when T=40 hours

```{r}
Time = c(40)
hard.new = as_tibble(Time)
hard.pred = predict(object = hard.lm, newdata = hard.new)
hard.pred
```

# Obtain point estimate of change in mean hardness in 1 hour

```{r}
delta_hard = reg_params$estimate[2]
se_delta_hard = reg_params$std_error[2]
```

The mean hardness change is `r delta_hard`, se = `r se_delta_hard`

